First off all, I want to thanks my advisor Prof. Carlos Henrique Costa Ribeiro for his precious support during this research. I am grateful for the patience and comprehension in the face of adversity that occurred throughout the program.

My special thanks to my co-advisor Cinara Guellner Ghedini not only for the outstanding support during this research and the encouragement during the program, but mainly for the friendship. 

I would also like to thank professors Carlos Henrique Quartucci Forster, Neusa Maria Franco de Oliveira and Lilian Berton for their valuable contribution as member of the Thesis Committee.

I would like to thank my family for supporting me in all my pursuits, specially to my brother Nataniel for the meticulous review of my thesis.

My special thanks to professors Anita Maria da Rocha Fernandes and Edson Tadeu Bez for the recommendation letters and for supporting me all the time. I would also like to thanks Jorge Destri Jr by its contribution by reviewing my master thesis and for the support during this research. Moreover, thanks to Diogo Andrei Benvenutti for the support in optimizing my source code in C++ and for encouragement. I am grateful to Caroline Mazzuco Furlan for the friendship and the support on mathematics aspects of the model. 

During this work, many friends and colleagues were very important. It is not possible to mention all, however some of these deserve to be highlighted. Thanks to my friends from Santa Catarina: Vitor Whays, Daniel Adriano, Rejane Araujo, Kerle Loffy, Ricardo Giordani and Leonardo Fenill. To my colleagues from ITA: Rene Esteves, Strauss Cunha, Rosana Teixeira, Mayara Moraes, Lucas Povoa, Jo{\~a}o Siles, Lais Siles and Paulo Diego Barbosa. In addition, thanks to my colleagues of Brazilian Air Force, especially to James de Castro Martins, Silvio Assun\c{c}ao and Ricardo Silva de Oliveira. 

I gratefully acknowledge CNPQ (Conselho Nacional de Desenvolvimento Cient{\'i}fico e Tecnol{\'o}gico) for the financial support.